---
title: "HMM"
output: html_document
---
```{r setup, include=FALSE}
library(tseries)
library(forecast)
library(rugarch)
library(depmixS4)
library(quantmod)
library(rmgarch)
library(ramify)
library(MTS)
library(vars)
library(depmixS4)
library(quantmod)
library(reticulate)
require(data.table)
```
install.packages("caret")
library(caret)
library(ggplot2)

```{r}
# import files
setwd("D:/My Projects/R 2021/Nifty")                    
files = list.files(pattern="ADANIPORTS.csv")
dataset = do.call(rbind, lapply(files, fread))
rm(files)

data <- as.data.frame(unclass(dataset))
data


```

```{r}
close=xts(data$Close, order.by=strptime(data$Date, format="%Y-%m-%d", tz=""))
#close
open=xts(data$Open, order.by=strptime(data$Date, format="%Y-%m-%d", tz=""))
#open
high=xts(data$High, order.by=strptime(data$Date, format="%Y-%m-%d", tz=""))
#high
low=xts(data$Low, order.by=strptime(data$Date, format="%Y-%m-%d", tz=""))
#low
#analyze data
plot(close,type='l')
diff_close=diff(close)
log_close=diff(log(close))
plot(log_close,type='l')



adf.test(na.remove(log_close))
acf(log_close)
Box.test(log_close,type = 'Ljung')
adf.test(log_close^2)
Box.test(log_close^2,type = 'Ljung')
summary(log_close)
summary(log_close^2)
plot(density(log_close),main="Daily Return")
xplot=seq(min(log_close),max(log_close),length=length(log_close))
yplot=dnorm(xplot,mean=mean(log_close),sd=sd(log_close))
lines(xplot,yplot,col="blue")
qqnorm(log_close,datax=T)
qqline(log_close,datax=T)
frac_change=(close-open)/open
frac_high=(high-open)/open
frac_low=(open-low)/open
frac_vec=as.matrix(cbind(frac_change,frac_high,frac_low))
ccm(frac_vec) 
```



```{r}

close=xts(data$Close, order.by=strptime(data$Date, format="%Y-%m-%d", tz=""))
#close
open=xts(data$Open, order.by=strptime(data$Date, format="%Y-%m-%d", tz=""))
#open
high=xts(data$High, order.by=strptime(data$Date, format="%Y-%m-%d", tz=""))
#high
low=xts(data$Low, order.by=strptime(data$Date, format="%Y-%m-%d", tz=""))
#low
volume=xts(data$Volume, order.by=strptime(data$Date, format="%Y-%m-%d", tz=""))
#volume
diff_close=diff(close)
log_close=diff(log(close))
length=length(log_close)
test1=log_close[1:(length-5)]
test_all=log_close[-1]
```

```{r}
frac_change=(close-open)/open
frac_high=(high-open)/open
frac_low=(open-low)/open
frac_vec=as.matrix(cbind(frac_change,frac_high,frac_low,volume))
#frac_vec
```
```{r}
close=xts(data$Close, order.by=strptime(data$Date, format="%Y-%m-%d", tz=""))
#close
open=xts(data$Open, order.by=strptime(data$Date, format="%Y-%m-%d", tz=""))
#open
high=xts(data$High, order.by=strptime(data$Date, format="%Y-%m-%d", tz=""))
#high
low=xts(data$Low, order.by=strptime(data$Date, format="%Y-%m-%d", tz=""))
#low
volume=xts(data$Volume, order.by=strptime(data$Date, format="%Y-%m-%d", tz=""))
#volume
```
```{r}
#HMM segmentation
close_x=xts(data$Close, order.by=strptime(data$Date, format="%Y-%m-%d", tz=""))
close_x=na.omit(cbind(close_x,diff(log(close_x))))
colnames(close_x)=c('close','return')
ATRindicator=ATR(data[,5:8],n=14)
ATR=ATRindicator[,2]
log_close=diff(log(close))
ModelData=data.frame(log_close,frac_vec)
ModelData
#ModelData=ModelData[-c(1:13),]
colnames(ModelData)=c("LogReturns","fra")



```


```{r}
##show fitted HMM states
length=length-1
model_data=data.frame(close_x[1:length])
log_close_test=log_close[1:length]
HMM = depmix(list(log_close_test~1), data = model_data, nstates = 3,family=list(gaussian()))
HMMfit=fit(HMM)
HMMpost=posterior(HMMfit)
plot.xts(close_x[1:length]$close,col='white')
points(close_x[1:length]$close[HMMpost$state==1],col='black')
points(close_x[1:length]$close[HMMpost$state==2],col='red')
points(close_x[1:length]$close[HMMpost$state==3],col='green')
#points(close_x[1:length]$close[HMMpost$state==4],col='blue')
```


```{r}
##transfer HMM states to dummy variables
log_close_x_pre=log_close[(length-5):length]
model_data=data.frame(close_x)
model_data_test=data.frame(close_x[(length-5):length])

close_x_test=close_x[1:(length-5)]
log_close_x_test=log_close[1:(length-5)]
model_data=data.frame(close_x_test)
log_close_x_pre=log_close[(length-5):length]
model_data_test=data.frame(close_x[(length-5):length])

HMM = depmix(list(log_close_x_test~1), data = model_data, nstates = 3, family=list(gaussian()))
HMMfit=fit(HMM)
HMMfit
HMMpost=posterior(HMMfit)
HMMpost
HMMfor=forwardbackward(setpars(depmix(list(log_close_x_pre~1),data=model_data_test,nstates = 3,family = list(gaussian())),getpars(HMMfit)))
HMMfor
pred_stat=argmax(HMMfor$alpha)
pred_stat
pred_stat=pred_stat[-1]
HMMpost_state=c(HMMpost$state,pred_stat)
HMMpost_state
```


```{r}
#HMM + gjrGARCH model to fit close price
dummy1=(HMMpost_state==1)
dummy2=(HMMpost_state==2)
dummy3=(HMMpost_state==3)
#dummy4=(HMMpost_state==4)
dummy=as.matrix(cbind(as.numeric(dummy1),as.numeric(dummy2),as.numeric(dummy3)))

test_dummy=dummy
#is.na(test_all)
#test_all
gspec=ugarchspec(variance.model=list(model='gjrGARCH',garchOrder=c(1,1)), mean.model=list(armaOrder=c(0,0),include.mean = TRUE,external.regressors=test_dummy[,1:2]), distribution.model = "sstd")
gspec
garch_fit=ugarchfit(spec=gspec,data=test_all,out.sample = 5)
garch_fit
garch_forcast=as.matrix(log_close[length-5])
for (i in (1:5))
{
garch_forcast=as.matrix(c(garch_forcast,garch_fit@model$pars[1,1]+garch_fit@model$pars[6,1]*test_dummy[length+i-5,1]+garch_fit@model$pars[7,1]*test_dummy[length+i-5,2]))
}
errors=as.numeric()
pred_price=as.numeric(close[length+1-5])
for (i in (1:5))
{
  pred=pred_price[length(pred_price)]*exp(garch_forcast[1+i])
  pred_price=c(pred_price,pred)
  errors=c(errors,(close[length+1+i-5]-pred)/close[length+1+i-5])
}

pred_price
errors
```
```{r}
compute_errors<-function(errors)
{
  avg_error=as.numeric()
  exp_error=as.numeric()
  for (i in (1:3))
  {
    total_avg=0
    total_exp=0
    for (j in (1:3))
    {
      total_avg=total_avg+abs(errors[i+j-1])
      total_exp=total_exp+abs(errors[i+j-1]*j)
    }
    avg_error=c(avg_error,total_avg/3)
    exp_error=c(exp_error,total_exp/6)
  } 
  return(cbind(avg_error,exp_error))
}
```

```
```{r}
ae_error=compute_errors(errors)
ae_error
```


```{r}
#HMM + VAR model to fit open/high/low
frac_change=(close-open)/open
frac_high=(high-open)/open
frac_low=(open-low)/open
frac_vec=as.matrix(cbind(frac_change,frac_high,frac_low))
frac_vec_test=frac_vec[2:(length-4),]

var_dummy=test_dummy[1:(length-5),]
var_dummy_test=test_dummy[(length-4):length,]
colnames(var_dummy)=c("exo1","exo2","exo3")
var_fit=VAR(frac_vec_test,exogen = var_dummy[,1:2])
var_pred=predict(var_fit,n.ahead=5,dumvar=var_dummy_test[,1:2])

open_pred=pred_price[-1]/(1+var_pred$fcst$frac_change[,1])
high_pred=open_pred*(1+var_pred$fcst$frac_high[,1])
low_pred=open_pred*(1-var_pred$fcst$frac_low[,1])
pred_price1=open_pred[-1]*(1+var_pred$fcst$frac_change[,1])
pred_price2=open_pred*(1-var_pred$fcst$frac_change[1])
#pred_price3=open_pred*(var_pred$fcst$frac_change[,1])
return=(open_pred/pred_price)-1
return[1]



#open_pred
#high_pred
#low_pred
pred_price1
#pred_price2
#pred_price3




plot(pred_price1)

```


```{r}
errors_open=(open[(length+1-4):(length+1)]-open_pred)/open[(length+1-4):(length+1)]
errors_high=(high[(length+1-4):(length+1)]-high_pred)/high[(length+1-4):(length+1)]
errors_low=(low[(length+1-4):(length+1)]-low_pred)/low[(length+1-4):(length+1)]
errors_open[-1:-2]
errors_high[-1:-2]
errors_low[-1:-2]
ae_errors_open=compute_errors(errors_open)
ae_errors_high=compute_errors(errors_high)
ae_errors_low=compute_errors(errors_low)
ae_errors_open
ae_errors_high
ae_errors_low
```

```{r}
#HMM + VAR model to fit open/high/low
frac_change=(close-open)/open
frac_high=(high-open)/open
frac_low=(open-low)/open
frac_vec=as.matrix(cbind(frac_change,frac_high,frac_low))
frac_vec_test=frac_vec[2:(length-30),]

var_dummy=test_dummy[1:(length-31),]
var_dummy_test=test_dummy[(length-30):length,]
colnames(var_dummy)=c("exo1","exo2","exo3")
var_fit=VAR(frac_vec_test,exogen = var_dummy[,1:2])
var_pred=predict(var_fit,n.ahead=31,dumvar=var_dummy_test[,1:2])
var_dummy_test
pred_price
open_pred=pred_price[-1]/(1+var_pred$fcst$frac_change[,1])
open_pred
close_pred=open_pred[-1]*(1+var_pred$fcst$frac_change[,1])
close_pred
high_pred=open_pred*(1+var_pred$fcst$frac_high[,1])
high_pred
low_pred=open_pred*(1-var_pred$fcst$frac_low[,1])
low_pred

plot(close_pred,type='o')
plot(high_pred,)

errors_open=(open[(length+1-31):(length+1)]-open_pred)/open[(length+1-31):(length+1)]
errors_open[1]
errors_high=(high[(length+1-31):(length+1)]-high_pred)/high[(length+1-31):(length+1)]
errors_high[1]
errors_low=(low[(length+1-31):(length+1)]-low_pred)/low[(length+1-31):(length+1)]
errors_low[1]
errors_open[-1:-2]
errors_high[-1:-2]
errors_low[-1:-2]
ae_errors_open=compute_errors(errors_open)
ae_errors_high=compute_errors(errors_high)
ae_errors_low=compute_errors(errors_low)
ae_errors_open
ae_errors_high
ae_errors_low

plot(pred_price)

```



```{r}

plot(close_pred,type='o')
#plot(open_pred,type='o')
#plot(high_pred,type='o')
#plot(low_pred,type='o')

lines(open_pred)
lines(high_pred)
lines(low_pred)



var_pred$fcst

```




```{r}
compute_errors<-function(errors)
{
  avg_error=as.numeric()
  exp_error=as.numeric()
  for (i in (1:3))
  {
    total_avg=0
    total_exp=0
    for (j in (1:3))
    {
      total_avg=total_avg+abs(errors[i+j-1])
      total_exp=total_exp+abs(errors[i+j-1]*j)
    }
    avg_error=c(avg_error,total_avg/3)
    exp_error=c(exp_error,total_exp/6)
  } 
  return(cbind(avg_error,exp_error))
}

ae_error=compute_errors(errors)
ae_error

#HMM + VAR model to fit open/high/low
frac_change=(close-open)/open
frac_high=(high-open)/open
frac_low=(open-low)/open
frac_vec=as.matrix(cbind(frac_change,frac_high,frac_low))
frac_vec_test=frac_vec[2:(length-4),]

var_dummy=test_dummy[1:(length-5),]
var_dummy_test=test_dummy[(length-4):length,]
colnames(var_dummy)=c("exo1","exo2","exo3")
var_fit=VAR(frac_vec_test,exogen = var_dummy[,1:2])
var_pred=predict(var_fit,n.ahead=5,dumvar=var_dummy_test[,1:2])
#pred_price
open_pred=pred_price[-1]/(1+var_pred$fcst$frac_change[,1])
open_pred
high_pred=open_pred*(1+var_pred$fcst$frac_high[,1])
high_pred
low_pred=open_pred*(1-var_pred$fcst$frac_low[,1])
low_pred
errors_open=(open[(length+1-4):(length+1)]-open_pred)/open[(length+1-4):(length+1)]
errors_high=(high[(length+1-4):(length+1)]-high_pred)/high[(length+1-4):(length+1)]
errors_low=(low[(length+1-4):(length+1)]-low_pred)/low[(length+1-4):(length+1)]
errors_open[-1:-2]
errors_high[-1:-2]
errors_low[-1:-2]
ae_errors_open=compute_errors(errors_open)
ae_errors_high=compute_errors(errors_high)
ae_errors_low=compute_errors(errors_low)
ae_errors_open

ae_errors_high
ae_errors_low
```